---
title: "Automated Machine Learning with H20 (I)"
author: "Tommy Weber"
---

### Importing Libraries

```{r}
#| message: false

library(tidyverse)
library(readxl)
library(skimr)
library(GGally)

```

---

### Loading Data

```{r}
#| message: false

employee_attrition_tbl <- read.csv("../../../employee_attrition.csv")

path_data_definitions <- "../../../data_definitions.xlsx"
definitions_raw_tbl   <- read_excel(path_data_definitions, sheet = 1, col_names = FALSE)
```

---

### Definition of data-wrangling functions

```{r}
#| code-fold: true

# Function to calculate attrition cost
calculate_attrition_cost <- function(

  # Employee
  n                    = 1,
  salary               = 80000,

  # Direct Costs
  separation_cost      = 500,
  vacancy_cost         = 10000,
  acquisition_cost     = 4900,
  placement_cost       = 3500,

  # Productivity Costs
  net_revenue_per_employee = 250000,
  workdays_per_year        = 240,
  workdays_position_open   = 40,
  workdays_onboarding      = 60,
  onboarding_efficiency    = 0.50

) {

  # Direct Costs
  direct_cost <- sum(separation_cost, vacancy_cost, acquisition_cost, placement_cost)

  # Lost Productivity Costs
  productivity_cost <- net_revenue_per_employee / workdays_per_year *
    (workdays_position_open + workdays_onboarding * onboarding_efficiency)

  # Savings of Salary & Benefits (Cost Reduction)
  salary_benefit_reduction <- salary / workdays_per_year * workdays_position_open

  # Estimated Turnover Per Employee
  cost_per_employee <- direct_cost + productivity_cost - salary_benefit_reduction

  # Total Cost of Employee Turnover
  total_cost <- n * cost_per_employee

  return(total_cost)

}


count_to_pct <- function(data, ..., col = n) {

  # capture the dots
  grouping_vars_expr <- quos(...)
  col_expr <- enquo(col)

  ret <- data %>%
    group_by(!!! grouping_vars_expr) %>%
    mutate(pct = (!! col_expr) / sum(!! col_expr)) %>%
    ungroup()

  return(ret)

}


assess_attrition <- function(data, attrition_col, attrition_value, baseline_pct) {

  attrition_col_expr <- enquo(attrition_col)

  data %>%
  
    # Use parenthesis () to give tidy eval evaluation priority
    filter((!! attrition_col_expr) %in% attrition_value) %>%
    arrange(desc(pct)) %>%
    mutate(
      # Function inputs in numeric format (e.g. baseline_pct = 0.088 don't require tidy eval)
      above_industry_avg = case_when(
        pct > baseline_pct ~ "Yes",
        TRUE ~ "No"
      )
    )
}
```


---

### Data-Wrangling

```{r}
dept_job_role_tbl <- employee_attrition_tbl %>%
  select(EmployeeNumber, Department, JobRole, PerformanceRating, Attrition)


dept_job_role_tbl %>%
  count(Department, JobRole, Attrition) %>%
  count_to_pct(Department, JobRole) %>%
  assess_attrition(Attrition, attrition_value = "Yes", baseline_pct = 0.088) %>%
  mutate(
    cost_of_attrition = calculate_attrition_cost(n = n, salary = 80000)
  )

```

---

### Definition of plot_ggpairs

```{r}
#| code-fold: true

plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) {
    
    color_expr <- enquo(color)
    
    if (rlang::quo_is_null(color_expr)) {
        
        g <- data %>%
            ggpairs(lower = "blank") 
        
    } else {
        
        color_name <- quo_name(color_expr)
        
        g <- data %>%
            ggpairs(mapping = aes_string(color = color_name), 
                    lower = "blank", 
                    legend = 1,
                    diag = list(continuous = wrap("densityDiag", 
                                                  alpha = density_alpha))) +
            theme(legend.position = "bottom")
    }
    
    return(g)
    
}
```

---

### Explore Features by Category

```{r}
#| message: false
#| warning: false

#   1. Descriptive features: age, gender, marital status 
employee_attrition_tbl %>%
    select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   2. Employment features: department, job role, job level
employee_attrition_tbl %>%
    select(Attrition, contains("employee"), contains("department"), contains("job")) %>%
    plot_ggpairs(Attrition) 
```

```{r}
#| message: false
#| warning: false

#   3. Compensation features: HourlyRate, MonthlyIncome, StockOptionLevel 
employee_attrition_tbl %>%
    select(Attrition, contains("income"), contains("rate"), contains("salary"), contains("stock")) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   4. Survey Results: Satisfaction level, WorkLifeBalance 
employee_attrition_tbl %>%
    select(Attrition, contains("satisfaction"), contains("life")) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   5. Performance Data: Job Involvment, Performance Rating
employee_attrition_tbl %>%
    select(Attrition, contains("performance"), contains("involvement")) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   6. Work-Life Features 
employee_attrition_tbl %>%
    select(Attrition, contains("overtime"), contains("travel")) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   7. Training and Education 
employee_attrition_tbl %>%
    select(Attrition, contains("training"), contains("education")) %>%
    plot_ggpairs(Attrition)
```

```{r}
#| message: false
#| warning: false

#   8. Time-Based Features: Years at company, years in current role
employee_attrition_tbl %>%
    select(Attrition, contains("years")) %>%
    plot_ggpairs(Attrition)
```

### Feature Investigation

#### 1. Compensation Features

What can you deduce about the interaction between Monthly Income and Attrition?

c) Those that are leaving have a lower Monthly Income

---

#### 2. Compensation Features

What can you deduce about the interaction between Percent Salary Hike and Attrition?

c) Those that are leaving have lower Percent Salary Hike

---

#### 3. Compensation Features

What can you deduce about the interaction between Stock Option Level and Attrition?

c) It's difficult to deduce anything based on the visualization

---

#### 4. Survey Results

What can you deduce about the interaction between Environment Satisfaction and Attrition?

a) A higher proportion of those leaving have a low environment satisfaction level

---

#### 5. Survey Results

What can you deduce about the interaction between Work Life Balance and Attrition

b) Those that are staying have a higher density of 2's and 3's

---

#### 6. Performance Data

What Can you deduce about the interaction between Job Involvement and Attrition?

a) Those that are leaving have a lower density of 3's and 4's

---

#### 7. Work-Life Features

What can you deduce about the interaction between Over Time and Attrition?

a) The proportion of those leaving that are working Over Time are high compared to those that are not leaving

---

#### 8. Training and Education

What can you deduce about the interaction between Training Times Last Year and Attrition

c) It's difficult to deduce anything based on the visualization

---

#### 9. Time-Based Features

What can you deduce about the interaction between Years At Company and Attrition

b) People that leave tend to have less working years at the company

---

#### 10. Time-Based Features

What can you deduce about the interaction between Years Since Last Promotion and Attrition?

b) Those that are leaving have fewer years since last promotion than those that are staying
